---
title: "dendrolenda"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

![check](https://github.com/michbur/dendrolenda/workflows/check/badge.svg)

Package for alignment of dendrogram plots with heatmap in ggplot2.

## Background

*dendrolenda* is a freely available software that allows users to visualize 
their data in the form of heat maps. It has an easy-to-use interface and makes 
alignments of dendrograms plots with heatmaps work like a charm. 

## Installation

*dendrolenda* is available on [github](https://github.com/michbur/dendrolenda). 
One can install the latest development version of the code using the following 
code:

```R
if("remotes" %in% rownames(installed.packages()) == FALSE) {
install.packages("remotes")
}
remotes::install_github("https://github.com/michbur/dendrolenda")
```

## Usage

```{r}
library(ggplot2)
library(dendrolenda)
data(example_data)
dendro_top <- plot_dendrogram(d1)
dendro_right <- plot_dendrogram(d2) + coord_flip()

# Plot aligned dendrograms with legend at the bottom
arrange_dendrogram(hm + theme_bw() + theme(legend.position = "bottom"),
dendro_top + theme_void(), dendro_right + theme_void())
```

